Four Models for Severe Weather Deaths

### 2017-05-10 paper.Rnw
Updated Methods, Results, and DIC for ZIP.

\setlength{\marginparwidth}{2cm}
\begin{document}
<<imports, cache=FALSE, echo=FALSE>>= 
read_chunk('zinb_sampler.R')
options(scipen=4, digits=3)
library('truncnorm')
@
<<dataset, echo=FALSE, cache=FALSE>>= 
load('small_storms.RData')
storm_details <- small_storms
@
\begin{center}
	\vspace{0.1cm}
	\textsc{\LARGE MATH 640 Final Project} \\[0.1cm]
	Jason Michaels (jam521), Niko Paulson (ndp32), Miranda Seitz-McLeese (mgs85) 
\end{center}
\section{Introduction}
\label{s:intro}
This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
The data set contains a variety of measures from severe weather events in the United States from 1996-2016. It was taken from the NOAA website \cite{dataset}.
For this project we focused on the deaths directly attributable to the event. In the interest of time we only modeled two types of severe weather: Tornadoes and Flash Floods.
Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
In this paper we compare four possible models for the deaths: The traditional Poisson and negative binomial distributions, as well as the zero inflated variant of each.

@@

\begin{table}
    \centering
    \caption{Results of taking 20,000 samples from the posterior distributions of $\lambda$ and p, compared between the two event types. The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses. The third does the same for p. All means and credible intervals are taken after discarding the first 10,000 samples as a burn-in.}
    \label{t:rZIP}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $\lambda$ & p  \\ \hline
    $Tornado$ & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
    $Flash Flood$ & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
    \end{tabular}
\end{table}

@@

\begin{table}
\centering
\caption{The following table shows the resulting DIC and WAIC for each of the four models on both the tornado data and the flash flood data. In the case of the zero inflated negative binomial model, WAIC was too computationally expensive, so it was not calculated.}
\label{t:evalresults}
\begin{tabular}{lcccc}
\toprule
&\multicolumn{2}{c}{Flash Flood}&\multicolumn{2}{c}{Tornado}\\
\cmidrule(r){2-3}\cmidrule(l){4-5}
Model & DIC &WAIC & DIC &WAIC\\
\midrule
Poisson & P FLOOD DIC& P FLOOD WAIC &P TORNADO DIC&P TORNADO WAIC\\
Negative Binomial & NB FLOOD DIC& NB FLOOD WAIC &NB TORNADO DIC&NB TORNADO WAIC\\
ZIP & ZIP FLOOD DIC& ZIP FLOOD WAIC &ZIP TORNADO DIC&ZIP TORNADO WAIC\\
ZINB & \Sexpr{flood.dic.zinb}& -- &\Sexpr{tornado.dic.zinb}&--\\
\bottomrule
\end{tabular}
\end{table}

>>>>

\setlength{\marginparwidth}{2cm}
\begin{document}
\SweaveOpts{concordance=TRUE}
<<imports, cache=FALSE, echo=FALSE>>= 
read_chunk('zinb_sampler.R')
options(scipen=4, digits=3)
library('truncnorm')
@
<<dataset, echo=FALSE, cache=FALSE>>= 
load('small_storms.RData')
storm_details <- small_storms
@
\begin{center}
	\vspace{0.1cm}
	\textsc{\LARGE MATH 640 Final Project} \\[0.1cm]
	Jason Michaels (jam521), Niko Paulson (ndp32), Miranda Seitz-McLeese (mgs85) 
\end{center}
\section{Introduction}
\label{s:intro}
This analysis will be done on a data set of a variety of measurements about severe weather in the United States. 
The data set contains a variety of measures from severe weather events in the United States from 1996-2016. It was taken from the NOAA website \cite{dataset}.
For this project we focused on the deaths directly attributable to the event. In the interest of time we only modeled two types of severe weather: Tornadoes and Flash Floods.
Understanding how and at what rate severe weather events become lethal in the United States has tremendous public health impacts.
In this paper we compare four possible models for the deaths: The traditional Poisson and negative binomial distributions, as well as the zero inflated variant of each.

@@

\begin{table}
    \centering
    \caption{Results of taking 20,000 samples from the posterior distributions of $\lambda$ and p, compared between the two event types. The second column relays the mean of the sample for $\lambda$, with the 95 percent credible interval in parentheses. The third does the same for p. All means and credible intervals are taken after discarding the first 10,000 samples as a burn-in.}
    \label{t:rZIP}
    \begin{tabular}{| l | l | l |}
    \hline
    Event Type & $\lambda$ & p  \\ \hline
    Tornado & 1.872 (1.030, 3.710) & 0.971 (0.968, 0.973) \\ \hline
    Flash Flood & 0.562 (0.293, 0.972) & 0.982 (0.981, 0.983) \\ \hline
    \end{tabular}
\end{table}

\includegraphics{figure/ZIP_Flood_Conv.pdf}

\noindent Figure 1. Cumulative mean plots for the posterior samples of the two parameters from the flood data. Plots for the tornado data are not pictured, but behave similarly.

\includegraphics{figure/ZIP_Density.pdf}

\noindent Figure 2. Posterior density plots for flood parameters (in blue), and tornado parameters (black).

@@

\begin{table}
\centering
\caption{The following table shows the resulting DIC and WAIC for each of the four models on both the tornado data and the flash flood data. In the case of the zero inflated negative binomial model, WAIC was too computationally expensive, so it was not calculated.}
\label{t:evalresults}
\begin{tabular}{lcccc}
\toprule
&\multicolumn{2}{c}{Flash Flood}&\multicolumn{2}{c}{Tornado}\\
\cmidrule(r){2-3}\cmidrule(l){4-5}
Model & DIC &WAIC & DIC &WAIC\\
\midrule
Poisson & P FLOOD DIC& P FLOOD WAIC &P TORNADO DIC&P TORNADO WAIC\\
Negative Binomial & NB FLOOD DIC& NB FLOOD WAIC &NB TORNADO DIC&NB TORNADO WAIC\\
ZIP & 3845.915& ZIP FLOOD WAIC &4759.998 &ZIP TORNADO WAIC\\
ZINB & \Sexpr{flood.dic.zinb}& -- &\Sexpr{tornado.dic.zinb}&--\\
\bottomrule
\end{tabular}
\end{table}
